---
title: "Homework 3"
author: "Diana Ballesteros Gonzalez (deg2163)"
date: "10/11/2018"
output: github_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

### **Problem 1** 
```{r setup_problem_1}
# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets")

## loading the library
library(p8105.datasets)
```


```{r load_brfss_smart2010}
## Load the data for brfss_smart2010
data("brfss_smart2010")
```

```{r brfss_smart_2010_df}
## cleaning the dataset 
brfss_smart2010_clean = 
   brfss_smart2010 %>% 
   janitor::clean_names() %>% 
   rename(state = locationabbr, county = locationdesc) %>% 
   filter(topic == "Overall Health" & (response == "Excellent" | 
                                       response == "Very good" |
                                       response == "Good" | response == "Fair" | 
                                       response == "Poor")) %>% 
   mutate(response = ordered(response, 
                             levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

```{r problem_1_questions}
## In 2002, which states were observed at 7 locations? Answer: CT, FL, and NC
brfss_smart2010_clean %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  distinct(county) %>% 
  count() %>% 
  filter(n == 7) 
```

```{r spaghetti_plot}
## Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.
brfss_smart2010_clean %>% 
  group_by(state, year) %>% 
  distinct(county) %>%
  summarize(count_county = n()) %>% 
  ggplot(aes(x = year, y = count_county, color = state)) + 
  geom_line() + 
  labs(
    title = "Number of locations in each state from 2002 to 2010",
    x = "Year",
    y = "Number of locations in each state",
    caption = "Data from the brfss_smart2010 package"
  ) + 
  viridis::scale_color_viridis(
    name = "States", 
    discrete = TRUE) + 
  theme_bw() 
```

```{r table_mean_sd}
## ## Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State. 
## SD I keep getting NA, need to convert from percentage to prop
brfss_smart2010_clean %>% 
  filter((year == 2002 | year == 2006 | year == 2010) & state == "NY" 
         & response == "Excellent") %>% 
  group_by(year) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  summarize(mean_excellent = mean(excellent),
            sd_excellent = sd(excellent, na.rm = TRUE)) %>% 
  mutate(mean_excellent = mean_excellent / 100, 
        sd_excellent = sd_excellent / 100)  %>% 
  knitr::kable(digits = 2)
```

```{r}
## For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
brfss_smart2010_clean %>% 
  group_by(year, state, response) %>% 
  summarize(mean_response_category = mean(data_value/100)) %>% 
  ggplot(aes(x = year, y = mean_response_category, color = state)) + 
  geom_line() + 
  facet_grid(~response) + 
  labs(
    title = "Distribution of state-level averages over time",
    x = "Year",
    y = "Average proportion in each response category",
    caption = "Data from the brfss_smart2010 dataset"
  ) + 
  viridis::scale_color_viridis(
    name = "State", 
    discrete = TRUE
  )
```

### **Problem 2**

```{r load_instacart_data}
## Load the data for instacart
data("instacart")
instacart = 
  instacart %>% 
  janitor::clean_names()
```

Description of the dataset: 
There are 15 variables in the dataset that include information including when the items were ordered, days since the customer's last order, product name, aisle the item was located at, product id, order id, whether the item was reorder, user's id. There are some variables that are not clearly defined including eval_set,order_dow. I am assuming that order_dow is referring to the day of the week the order was placed from 0-6 with 0 being Saturday according to some online research since those are the days with the most instacart orders. The dataset is structured so that each row is a different item from an order that was placed (i.e., 1,384,617 items were ordered via instacart). 


```{r aisle_calculation}
## How many aisles are there?
instacart %>% 
 distinct(aisle_id) %>%
 count()
```

There are `r instacart %>% distinct(aisle_id) %>% count()` aisles. 

```{r aisle_most_items}
## Which aisles are the most items ordered from?
instacart %>% 
  group_by(aisle) %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>% 
  head() %>% 
  knitr::kable()
```

```{r plot_aisles_most_items}
instacart %>% 
  group_by(aisle) %>% 
  count(aisle) %>%
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(
    title = "Number of items ordered in each aisle",
    x = "Aisle",
    y = "Number of items ordered per aisle",
    caption = "Data from the instacart dataset"
  ) + 
 theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r most_pop_item}
## Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
instacart %>% 
  filter(aisle == "baking ingredients" | 
         aisle == "dog food care" | 
         aisle == "packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  summarize(most_popular_item = max(product_name)) %>%
  knitr::kable()
```

```{r}
## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
instacart %>% 
  filter(product_name == "Pink Lady Apples" | 
         product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_order_hour_of_day = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_order_hour_of_day) %>% 
  rename("Product Name" = "product_name", "Saturday" = "0", 
         "Sunday" = "1", "Monday" = "2", "Tuesday" = "3", 
         "Wednesday" = "4", "Thursday" = "5", "Friday" = "6") %>% 
  knitr::kable(digits = 2)
```

### **Problem 3** 
```{r load_nyoaa}
## Load the data for instacart
data("ny_noaa")
ny_noaa = ny_noaa 
```

Description of the dataset: 
There are 7 vairables in this dataset with 2,595,176 observations (i.e., number of row). Each row provides information on the precipitation, max and minimum temperature on any given day. Each dataset is : 
id = station identification code
prcp = Precipitation (tenths of mm)
snow = Snowfall (mm)
snwd = Snow depth (mm)
tmax = Maximum temperature (tenths of degrees C)
tmin = Minimum temperature (tenths of degrees C)
  
```{r}
## Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 

## Separated date into year, month, and day; converted month to list the, converted the units for prcp to inches, tmin/tmax from tenths of degrees C to degrees C
ny_noaa_clean = 
  ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year","month","day"), sep = "-") %>% 
  mutate(month = month.name[as.numeric(month)],
         tmin = as.numeric(tmin, na.rm = TRUE),
         tmax = as.numeric(tmax, na.rm = TRUE), 
         prcp = prcp * 10 / 25.4,
         tmin = tmin / 10,  
         tmax = tmax / 10
         ) 
```

```{r}
## For snowfall, what are the most commonly observed values? Why?
ny_noaa_clean %>%
  group_by(snow) %>% 
  count(snow) %>% 
  summarize(most_common_observed = max(n, na.rm = TRUE)) %>% 
  arrange(desc(most_common_observed)) %>% 
  top_n(1)
```

```{r}
## Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

ny_noaa_clean %>% 
  filter(month == "January" | month == "July") %>% 
  group_by(year, month) %>% 
  mutate(average_temp_per_month = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = average_temp_per_month)) + 
  geom_point() + 
  facet_grid(~month) + 
  labs(
    title = "Average Max Temperature in January and July in Each Station Across Years",
    x = "Year",
    y = "Average Max Temperature",
    caption = "Data from the nyoaa dataset"
  ) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) 
```

```{r}
## Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

## Load the patchwork library
library(patchwork)

## Install the hexbin package and load it 
library(hexbin)

## Create tmax vs tmin plot 
ny_noaa_tmin_tmax = 
  ny_noaa_clean %>%
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex() + 
  labs(
    title = "Minimum vs Maximum Temperature",
    x = "Minimum Temperature",
    y = "Max Temperature") + 
  theme(legend.position = "none")

## Create plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
ny_noaa_snow = 
  ny_noaa_clean %>%
  group_by(year) %>% 
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x = year, y = snow)) + 
  geom_boxplot() + 
   labs(
    title = "Distribution of snowfall values greater than 0 and less than 100 by year",
    x = "Year",
    y = "Amount of Snowfall in mm") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) 
  
## Using patchwork, I combined both plots 
ny_noaa_tmin_tmax + ny_noaa_snow

```

